The Discounted Cash Flow model can be summarized by the following formula \parencite{brealey2020principles}:

\begin{equation} \label{eq:dcf_formula}
    \sum_{n=1}^{N}\frac{FCFF_n}{(1+r)^n} + {\frac{TV}{(1+r)^N}}
\end{equation}

where \(FCFF_n\) is the free cash flow to the firm for the \(n^{th}\) year , \(r\) is the rate used to discount the cash flows, \(TV\) is the terminal value and \(N\) is the number of periods (years) in the valuation.

The fundamental premise of the DCF model is that the intrinsic value of any cash-generating asset, such as a company, equals the present value of its expected future free cash flows. The formula captures this by summing the discounted values of projected cash flows over the forecast period.

The second component, the discounted terminal value, accounts for the company's value beyond this finite forecast horizon. Projecting individual cash flows indefinitely is impractical due to diminishing forecast accuracy over longer time frames. Consequently, the model typically assumes that after year \(N\), the company enters a "steady state" of perpetual, stable growth. This terminal value, often calculated based on the final projected cash flow and an assumed long-term growth rate (usually constrained by the nominal growth rate of the broader economy), represents the value of all cash flows generated after the last year of the forecasting period. This future value \((TV)\) is then discounted back to the present using the same discount rate.

The sum of these two components – the present value of each period cash flows and the present value of the terminal value – yields the estimated intrinsic value of the company's operating assets.

The following chapters will focus on each of the core inputs to this equation (\(FCFF_n,\space TV,\space N,\space r\)) providing a framework for their estimation and utilization within the valuation model.

\input{parts/part_1/forecasting_period}

\input{parts/part_1/dicount_rate}

\input{parts/part_1/fcff}

\input{parts/part_1/terminal_value}

\input{parts/part_1/valuation_adjustments}

Subtracting the estimated value of management and employee stock options represents the final major adjustment before arriving at the total intrinsic value of common equity available to current shareholders. This value is then divided by the current number of outstanding common shares to calculate the intrinsic value per share.