\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{Introduction}
\chapter*{Introduction}

On November 30, 2022, the general public was introduced to Artificial Intelligence (AI) at an unprecedented scale. This date marked the release of ChatGPT by \textcite{openai2022chatgpt}, inaugurating the era of the mass-market AI chatbot. While it represents the latest in a long lineage of great innovations in human history, it quickly revealed profound and significant differences from its predecessors.

The adoption rate was staggering—reaching one million users in just five days—a metric that immediately underscored the sheer scale and velocity of this technological shift. For comparison, it took 74 days for the iPhone to reach its one-millionth sale in 2007 \parencite{Apple2007Million}.

Furthermore, the most crucial difference between AI and past paradigm-shifting innovations (such as electricity or the automobile) lies in its technological nature: AI models are highly intangible and conceal an enormous degree of underlying complexity. The fields of Machine Learning, Artificial Intelligence, and Computer Science that power these products are among the vastest and most complex subjects in human knowledge. Crucially, managing and understanding such innovations requires deep specialization, implying an ever-increasing need for industry specialists as opposed to generalists.

Earlier that same year, on February 24, 2022, Russia invaded Ukraine, igniting a war at the heart of Europe that continues to reshape global geopolitics years later. Although \textcite{fukuyama1989end} had famously posited \textit{The End of History?} decades prior, and the idea of a stagnating, peaceful historical progression had taken root in much of the West during the 2000s, that day in February served as a brutal reminder to the Old Continent that history is far from static.

\section*{The Cross-Dimensionality of Valuation}

As these two monumental developments demonstrate, human history is neither complete nor simple. We are currently navigating a world characterized by unprecedented speed and complexity, subject to defining macroeconomic and geopolitical events that are faster and deeper than ever before. This dynamic environment is the backdrop against which all social and economic structures, and by extension all corporate entities, operate.

A central thesis of this work is that in today's world, valuing a company can no longer be treated as the isolated prerogative of finance or economics departments. As the following chapters will demonstrate, executing a robust valuation requires a holistic, cross-dimensional understanding of both the microeconomic environment and the broader global background in which a firm operates.

Legendary investors such as Warren Buffett, Charlie Munger, and Peter Lynch have long advocated for this approach, famously advising to \textit{``Buy what you know.''} This thesis aims to formalize why they were right, examine if their maxim remains applicable today, and define the constraints under which it operates.

As will be detailed in Chapter \ref{chap:cash_flow}, executing a Discounted Cash Flow (DCF) valuation inevitably requires estimating key metrics for a company: future revenues, operating margins, investment needs, and growth rates. While forecasting is inherently difficult and highly susceptible to human error and cognitive biases \parencite{kahneman2011thinking}, it is widely accepted that holding superior, specialized information results in more accurate estimates, provided those biases are properly managed.

As technology becomes increasingly embedded within corporate operations, and as it demands greater specialization to be utilized and understood, the pool of analysts capable of truly understanding modern companies naturally shrinks. The legendary investors mentioned above made their fortunes predominantly with relatively simple, easy-to-understand businesses (such as Coca-Cola or See's Candies). One could argue that, at the time of their purchase, these were the ``tech'' stocks of their era. Yet, even under this argument, one fact remains crystal clear: the sheer volume and complexity of technology embedded in a modern company like Apple is exponentially greater than that found in a traditional consumer goods company like Coca-Cola \parencite{broekel2025long}.

If corporate complexity is continuously increasing, and our valuation frameworks demand that we invest where our knowledge lies, it follows that properly valuing a modern company requires an analyst to possess a vast, deep understanding of the industry, the product, and the specific business model. Thus, modern valuation is inherently \textit{cross-dimensional and interdisciplinary}.

\paragraph{The Alphabet Inc. (Google) AI Story}

To illustrate this necessary cross-dimensionality, the strategic positioning of Alphabet Inc. (Google's parent company) within the AI industry serves as a compelling example. This company provides a clear demonstration of why modern corporate valuation cannot be reduced to a mere accounting exercise.

As explicitly stated in its latest Form 10-K \parencite{alphabet202410k}, \textit{``Alphabet is a collection of businesses.''} Among these subsidiaries, Google stands out as a foundational player in the global AI race, maintaining a presence across virtually every layer of the AI technology stack. It is here that the interdisciplinary requirements of valuation become apparent.

Observers outside the tech industry may view AI simply as an ``app that answers questions.'' While true on the surface, this represents only a fraction of the reality. At its core, generative AI relies on immensely complex statistical models. Simplistically, these models are massive mathematical equations containing trillions of parameters (or ``knobs''). The objective is to adjust these trillions of parameters so that the model's output (its answer) is as accurate as possible.

Because no human can manually adjust trillions of parameters, this task is outsourced to highly specialized computers. Traditional processors (CPUs) execute computations sequentially. However, adjusting AI parameters requires massive parallelization, a task for which Graphics Processing Units (GPUs) are uniquely suited. Yet, even GPUs are general-purpose parallel processors. Because the mathematical operations required for AI are highly specific (primarily matrix multiplications), the industry has developed Application-Specific Integrated Circuits (ASICs)—processors custom-tailored to excel exclusively at these specific parallel operations.

The computationally intensive process of finding the correct parameters is known as \textit{training}. Once trained, the model enters the \textit{inference} phase, where it is actually utilized by end-users to generate responses.

Zooming out, the AI industry can be segmented into several key pillars: foundational models, training hardware, and inference infrastructure. At every stage of this value chain, a multitude of companies interact, much like any traditional economic sector. For a valuation analyst, understanding this architecture is paramount, as a company's position within this stack will dictate its future margins, pricing power, supply chain vulnerabilities, and growth trajectory.

Alphabet's integration across this entire stack is profound. Starting with the models, Alphabet acquired DeepMind in 2014 and merged it with the Brain team from Google Research in 2023 \parencite{pichai2023merger}. As noted in the official merger announcement, this brought together the teams responsible for landmark achievements such as AlphaGo, AlphaFold, and TensorFlow. Crucially, it was Alphabet researchers who published the seminal paper \textit{``Attention Is All You Need''} \parencite{vaswani2017transformer}, which introduced the Transformer architecture that serves as the foundation for modern Large Language Models (LLMs).

Beyond software models, Alphabet is extremely active in AI hardware. Its proprietary ASICs, known as Tensor Processing Units (TPUs), are currently in their seventh generation. Notably, the Gemini 3.0 Pro and Gemini 3.1 Pro models were trained entirely on Google's own TPUs \parencite{gemini3pro_model_card}. This vertical integration is a massive value driver: by utilizing its own hardware, Alphabet reduces its dependency on third-party suppliers like NVIDIA, retaining the margin that would otherwise be paid externally and significantly lowering the costs of both training and inference. Furthermore, successfully training a flagship model like Gemini on TPUs served as a vital proof-of-concept for the broader industry, demonstrating that NVIDIA is not the only viable hardware provider. Consequently, companies like Meta have actively secured deals to utilize Google's TPUs for their own AI workloads \parencite{the_information2026meta_tpu}.

At the application layer, Alphabet has integrated the Gemini model into its consumer-facing apps and services, securing its position as a top player in end-user AI deployment.

For an analyst modeling Alphabet's DCF, understanding this full-stack approach is critical. It implies that the Total Cost of Ownership (TCO) for Gemini models should be structurally lower than that of non-integrated competitors. This cost advantage could translate into the ability to launch aggressive discounting or marketing campaigns to capture market share while maintaining profitability.

However, forecasting requires deep technical nuance. While it is true that Meta is utilizing TPUs, a naive analyst might blindly forecast that the entire industry will abandon NVIDIA. This would demonstrate a fundamental misunderstanding of the sector. The true ``moat'' protecting NVIDIA's dominance is not just its hardware, but its software stack. Modern chips ``speak'' machine language, which humans cannot directly manage. NVIDIA's proprietary programming platform, CUDA, has become the absolute industry standard. Because Google's TPUs operate on a different framework (JAX), transitioning away from NVIDIA requires immense software engineering resources. Thus, while well-funded teams at Meta can bridge this gap, it is analytically unsafe to assume the broader market will easily follow suit.

Finally, Alphabet monetizes AI workloads through Google Cloud Platform (GCP), selling a vast suite of services for developers to build, manage, and scale AI models. As \textcite{james2023leading} noted: \textit{``More than half of all funded gen AI startups are Google Cloud customers, including 70\% of gen AI `unicorns'.''} This high market penetration implies two key valuation drivers: first, Google profits from the broader success of the AI industry regardless of whose models win (mitigating the risk of a zero-sum game); second, capturing these startups early may spark ancillary growth across GCP's non-AI enterprise services.

This brief exploration of Alphabet illustrates why modern valuation requires experts who grasp both the financial mechanics of a DCF and the underlying technological constraints the company must navigate. Modeling key financial metrics has become so tightly intertwined with business characteristics that the analyst must possess deep, holistic industry expertise. As the reader progresses to \textit{Part 2} of this thesis, this logic will return forcefully. In a world where companies specialize in ever-growing technological complexities, deep knowledge is a non-negotiable prerequisite for accurate valuation.

\section*{Valuation in a Changing World Order}

If the OpenAI narrative highlights the dizzying pace of technological innovation, the 2022 Russian invasion of Ukraine served as a stark reminder that economic activity does not occur in a vacuum. Rather, it is deeply embedded within an interconnected, and increasingly fragmented, geopolitical landscape. The invasion, alongside escalating global tensions involving the United States, Venezuela, and the Middle East, underscored the fragility of the post-WWII international order and proved that history is continuously evolving.

Crucially, these events demonstrated how the global financial system can be effectively "weaponized" during periods of severe conflict. Following the invasion, the Russian Central Bank was forced to halt ruble trading in foreign exchange markets, while Western nations imposed unprecedented coordinated sanctions. These measures included the outright prohibition of trading Russian securities for Western investors and the expulsion of key Russian institutions from SWIFT—the highly centralized messaging bedrock of international finance.

These developments compel a fundamental reexamination of what valuation truly represents and the assumptions upon which it relies. From a purely theoretical standpoint, the existence of a secondary trading market does not inherently alter the cash-generating capacity of an operating asset. Private businesses, for instance, obviously hold immense intrinsic value despite lacking a public exchange. Traditional corporate finance often handles this discrepancy by applying an "illiquidity discount." While it is common sense that a highly liquid, publicly traded stock like Apple carries less friction risk than a privately owned enterprise, mechanically penalizing a company solely for lacking a public ticker oversimplifies the modern risk landscape.

The ultimate objective of any valuation is to inform an actionable investment decision: buy, hold, or sell. Executing these decisions requires functioning, centralized financial plumbing—wiring funds, clearing trades, and moving capital across borders. The weaponization of these channels introduces a binary "realization risk." If an investor cannot legally or practically access a market, or repatriate their capital due to geopolitical sanctions, the theoretical present value of the asset's cash flows becomes largely irrelevant.

Consequently, estimating the true risk of an asset requires moving beyond crude macroeconomic indicators (such as inflation or debt-to-GDP ratios). Analysts must comprehensively evaluate a country's geopolitical alliances, its integration into the global political economy, its domestic stability, and the strategic trajectory of its ruling elite. We can no longer ignore the political alignment of the jurisdiction in which a company operates or its relationship with the state apparatus.

The key insight here is that cross-border financial investments must increasingly be analyzed through the lens of Foreign Direct Investment (FDI) in real, physical assets. When a multinational corporation constructs a factory in a foreign jurisdiction, it operates under the assumption that the physical asset is essentially captive; it cannot simply be dismantled and brought back home overnight. The investment will only yield returns if the local political and economic environment remains stable enough for the factory to operate profitably. In the modern geopolitical landscape, portfolio investments share far more of this captive, locally-dependent risk than market participants often acknowledge. A stock certificate may be theoretically liquid, but if sanctions, capital controls, or institutional collapse sever the financial plumbing, that paper claim becomes just as stranded as a physical factory.

This leads to a profound realization: intrinsic value is highly context-dependent. A Russian corporation generating robust, predictable cash flows might possess an intrinsic value of \$100 billion to a domestic investor residing in Moscow, but an effective value of zero to a sanctioned US citizen. The inputs—the cash flows and the growth rates—remain identical, yet the institutional barrier dictates a completely divergent outcome. Far from undermining the scientific validity of valuation, this dynamic simply reinforces a core tenet of economics: value is inherently subjective, much like a cup of water commands a vastly different price in a restaurant versus the middle of a desert. The key insight for the modern analyst is not to avoid valuing such companies altogether, but to dynamically adjust the framework and the resulting output for different audiences and geopolitical scenarios.

Recent empirical research strongly supports this jurisdictional divergence in valuation. As demonstrated by \textcite{lei2025democracy} in \textit{Democracy and Stock Market Returns}, countries with more inclusive, democratic institutions command higher equity valuations and deliver superior returns. Crucially, the researchers found that standard asset pricing models, such as the Capital Asset Pricing Model (CAPM), failed to fully explain this return differential, even after accounting for diversifiable country-specific risks. This reinforces the core thesis of this work: two mathematically identical companies will command fundamentally different valuations based solely on the political and institutional quality of the jurisdictions in which they operate.


\section*{A Framework for Valuation}

As time passes and environments evolve, the quest to understand, quantify, and reason about economic value has accompanied humanity throughout the centuries. Within the vast ocean of economic study, the discipline of asset and company valuation must continuously be reconsidered and adjusted to reflect the realities of the modern world.

The structure of this thesis reflects the necessity of reconciling the strict, quantitative approach of Corporate Finance with the prolific and ever-growing insights of institutional economics, macroeconomics, economic sociology, and history. To achieve this, the thesis is structured into two distinct but deeply interconnected parts.

\subsection*{Part I: Building the Model}
The first part, \textit{Building the Model}, establishes the mathematical and theoretical framework of the thesis. The core philosophy of valuation is straightforward: the intrinsic value of any asset is simply the sum of all the cash it is expected to generate in the future, discounted back to its present value today. The bedrock upon which this framework is built is the time value of money—the principle that, assuming positive inflation and opportunity costs, a dollar today is worth more than a dollar tomorrow. Discounting is the mathematical mechanism used to assign a present-day value to future cash flows.

This foundational concept leads directly to the core questions of a Discounted Cash Flow (DCF) model: How far into the future should cash flows be explicitly forecast? What is the appropriate risk-adjusted rate at which to discount them? How do we accurately estimate the cash the business will generate? And how do we capture the value of the firm beyond our capacity to forecast accurately? Chapters \ref{chap:forecasting_period} through \ref{chap:valuation_adjustments} provide the theoretical answers and practical nuances to each of these questions.

\textbf{Chapter \ref{chap:forecasting_period}} addresses the time horizon of the valuation, focusing on determining the optimal length of the explicit forecast period based on the company's specific characteristics. The core premise is that all companies eventually enter a ``steady-state''—a mature phase where operations become highly predictable. The forecasting period models the transition from the company's current state until it reaches this predictable maturity.

\textbf{Chapter \ref{chap:discount_rate}} explains the derivation of the proper discount rate, the critical lever used to adjust future cash flows for risk. This factor must encapsulate inflationary risks, broad market volatility, company-specific systemic risks, country and political instability, and default risk. The chapter explores several methodologies to arrive at this final rate, weighing the advantages and disadvantages of each.

\textbf{Chapter \ref{chap:cash_flow}} tackles the complex task of estimating the future cash flow stream prior to discounting. It establishes a framework for estimating the true cash generated by operations that is theoretically available to capital providers, after accounting for all operating expenses, taxes, and the reinvestment (Capital Expenditures and Working Capital) required to sustain and grow the business. For conglomerate entities, this involves a granular Sum of the Parts (SOTP) approach, ensuring that revenue, margin, and reinvestment assumptions are tailored to the specific end-markets of each business segment.

\textbf{Chapter \ref{chap:terminal_value}} analyzes the mechanics of capturing value beyond the explicit forecasting period. At the end of the forecast horizon, a company faces two primary theoretical paths: liquidation or continued perpetual operation. This chapter explores how to mathematically model both scenarios to capture the firm's terminal value.

\textbf{Chapter \ref{chap:valuation_adjustments}} concludes the first part by detailing the final adjustments required to bridge the gap between the value of the firm's operating assets and the true intrinsic value of its common equity. This includes accounting for cash equivalents, outstanding debt, cross-holdings, and the dilutive impact of management stock options.

\subsection*{Part II: Model Response to Macroeconomic Shocks}

The second part, \textit{Model Response to Macroeconomic Shocks}, attempts the audacious task of grounding the sterile mathematical model built in Part I within the volatile, real-world environment in which companies actually operate. The core philosophy of this thesis is that corporate valuation must never be conducted in a vacuum; it must be tightly integrated with macroeconomics, monetary policy, developmental economics, and history.

These broader fields provide the frameworks necessary to understand the arena in which economic activity takes place. If macroeconomic and monetary forces represent the ``laws of physics'' governing supply, demand, and capital flows, then institutional and developmental economics provide the ``rules of the game''—the legal and political frameworks overlaid upon those physical laws. History and sociology, in turn, shed light on how human actors navigate, react to, and alter those rules over time.

Unlike physical systems, the economic ecosystem is mutable. Its parameters change constantly, whether through shifts in top-down trade policies and tariffs, or bottom-up changes in consumer preferences and technological adoption. The second part of this thesis provides a structured framework for anticipating how these macro-level shocks cascade down to impact the specific micro-level inputs (TAM, margins, reinvestment, and risk) of the DCF model.

\textbf{Chapter \ref{chap:interest_rates}} tackles one of the most powerful forces in any economy: the Central Bank. We examine the mechanisms of monetary policy and observe their ripple effects throughout the economy, investigating both absolute changes in interest rates and relative shifts in the yield curve. The chapter explores how these monetary shifts impact distinct industry sectors and ultimately flow directly into the discount rate calculations established in Chapter \ref{chap:discount_rate}.

\textbf{Chapter \ref{chap:aggregate_demand}} bridges the gap between macroeconomic theory and corporate forecasting by dissecting Aggregate Demand. By deeply analyzing Consumption, Investment, and Government Spending cycles, this chapter establishes a framework to connect top-down GDP fluctuations directly to the bottom-up cash flow and revenue estimates required in Chapter \ref{chap:cash_flow}.

\textbf{Chapter \ref{chap:international_trade}} explores the complexities of the globalized economy. It details how exchange rate fluctuations and international trade flows impact domestic and multinational firms. The chapter then focuses heavily on the disruptive effects of tariffs and trade restrictions, illustrating how assumptions regarding market access, competitive margins, and growth must be dynamically revised in the face of protectionism.

\textbf{Chapter \ref{ch:tech_change}} dives into the pervasive impacts of technological change on valuation. It highlights how the process of ``creative destruction'' forces the continuous reassessment of DCF inputs, explaining how the financial consequences of disruption differ vastly for the innovator versus the incumbent.

\textbf{Chapter \ref{chap:political_risk}} concludes the thesis by linking institutional economic frameworks—specifically the Nobel-winning AJR framework regarding inclusive versus extractive institutions—directly to the valuation model. We review how the historical evolution of a country's institutions determines its baseline stability, and how analysts must map this institutional quality directly onto their assessments of systemic risk, reinvestment incentives, and terminal growth constraints.